% =============================================================================
% paper_orbitalif.tex
%
% ORBITALIF: Orbital Attention-enhanced Leaky Integrate-and-Fire framework
%            for decentralized onboard federated optical cloud removal.
%
% NEW DRAFT (user-supplied abstract + introduction).  This file is intended
% to be the starting point for the user's own further editing; method,
% experiments, and conclusion sections are NOT included here yet.
%
% --- Reference verification convention --------------------------------------
%   Every \bibitem below is tagged in a comment with one of:
%     [VERIFIED]      — author / title / venue / year cross-checked against
%                       the actual primary source; OK to cite as-is.
%     [USER VERIFY]   — paper exists but exact bibliographic fields (page
%                       numbers, volume, doi) need user cross-check.
%     [PLACEHOLDER]   — generic stub; user must supply a real reference
%                       before submission.
% --- IMPORTANT --------------------------------------------------------------
%   Do NOT submit this paper while any \bibitem is still tagged [PLACEHOLDER]
%   or while [USER VERIFY] entries have not been confirmed against the
%   published or arXiv version.
% =============================================================================

\documentclass[conference]{IEEEtran}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{caption}
\hypersetup{
  colorlinks   = true,
  linkcolor    = blue,
  citecolor    = blue,
  urlcolor     = blue,
}

% --- Useful macros ---
\newcommand{\E}{\mathbb{E}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\Dir}{\mathrm{Dir}}
\newcommand{\OrbitALIF}{\textsc{OrbitALIF}}

\graphicspath{{figures/}}

\begin{document}

% --- Title block ---
\title{\OrbitALIF: Orbital Attention-enhanced Leaky Integrate-and-Fire
       for Decentralized Onboard Federated Optical Cloud Removal}

\author{
  \IEEEauthorblockN{(authors anonymised for review)}\\
  \IEEEauthorblockA{(institutions anonymised for review)}
}

\maketitle

% =====================================================================
% Abstract
% =====================================================================
\begin{abstract}
Satellite networks can acquire massive space information with advanced
remote sensing technologies, providing critical support for real-time
applications such as natural disaster monitoring. Emerging Space
Computing Power Networks provide a new architecture for coordinating
onboard computing resources, enabling satellites to perform intelligent
in-orbit processing collaboratively. Nevertheless, optical cloud
removal is a high-fidelity dense restoration task, and existing
artificial-neural-network-based models usually incur high computation
and energy costs, making them difficult to sustain under the limited
power supply and thermal conditions of satellite platforms.
To tackle this issue, we propose \OrbitALIF{}, which exploits the
sparse event-driven computation of spiking neural networks to reduce
inference energy, while enhancing pixel-level restoration capability
through attention- and receptive-field-enhanced modules and
architecture.
To enable onboard model training, we further formulate optical cloud
removal as a decentralized onboard federated learning problem over
satellite constellations. To the best of our knowledge, this is the
first work to study this problem. To address source-level heterogeneity
in satellite observations, \OrbitALIF{} preserves local
batch-normalization statistics to mitigate feature shift and exchanges
shared parameters through intra-plane aggregation and inter-plane
Gossip mixing. Experiments show that the $2.30$\,M onboard deployment
variant improves cloud-removal quality while reducing per-image
inference energy by up to $72.3\!\times$. The scaled variant achieves
SOTA performance on image deraining and competitive near-SOTA
performance on cloud removal. These results demonstrate a new pathway
toward low-energy, decentralized, and satellite-native optical
remote-sensing restoration.
\end{abstract}

\begin{IEEEkeywords}
Spiking neural networks, optical cloud removal, decentralized
federated learning, onboard satellite computing.
\end{IEEEkeywords}

% =====================================================================
% Section I: Introduction
% =====================================================================
\section{Introduction}\label{sec:intro}

Low-Earth-orbit optical remote-sensing satellite networks
continuously acquire large-scale space information, providing critical
data support for real-time applications such as natural disaster
monitoring, emergency response, agricultural observation, environmental
assessment, and defense surveillance~\cite{ref:earthobs1,ref:earthobs2}.
However, optical remote-sensing imagery is highly vulnerable to cloud
contamination, which degrades image usability and affects downstream
tasks such as object detection, land-cover classification, and change
detection. Conventional processing pipelines usually follow a
``satellite acquisition--raw data downlink--ground-server processing''
paradigm, where cloud-contaminated images are first transmitted to
ground stations and then restored and analyzed on ground servers. As
satellite constellations scale up and imaging resolution increases,
this paradigm faces an increasingly severe timeliness bottleneck:
visible contact windows between satellites and ground stations are
limited, downlink bandwidth is constrained, and ground-side processing
further introduces additional latency. Existing studies on onboard
federated learning have also pointed out that satellite data
transmission is a critical bottleneck for Earth-observation
applications, directly limiting real-time space information
services~\cite{ref:onboardfl-bottleneck}.

Space Computing Power Networks (Space-CPN) provide a promising
architecture to alleviate the above bottleneck. Instead of treating
satellites merely as data acquisition and forwarding nodes, Space-CPN
emphasizes the coordinated scheduling of communication, computation,
and intelligent processing resources in space networks, enabling
satellites to undertake part of the in-orbit data processing
workload~\cite{ref:spacecpn1}. For optical cloud removal, this paradigm
is particularly meaningful. Cloud contamination is a fundamental
problem that must be addressed before optical remote-sensing data can
be reliably used by downstream tasks. If cloud removal can be performed
onboard, the downlink of low-quality raw observations can be reduced,
the response speed of time-critical tasks can be improved, and cleaner
inputs can be provided for subsequent interpretation. Recent studies on
Space-CPN and in-orbit computing also regard distributed onboard
computation as an important direction for future space information
processing~\cite{ref:spacecpn1}.

Nevertheless, onboard optical cloud removal is not a simple transfer
of ground-side models to satellites. First, optical cloud removal is a
high-fidelity dense image restoration task. It requires pixel-level
recovery of cloud boundaries, thin-cloud textures, ground structures,
and high-frequency details. Existing artificial neural network
(ANN)-based cloud-removal models typically rely on continuous-valued
activations and dense multiply-accumulate operations, resulting in
considerable computation and energy costs. In particular, recent
diffusion-based cloud-removal methods have achieved strong performance
on high-resolution datasets such as CUHK-CR through progressive texture
recovery, but their iterative inference process is inherently in
tension with the limited power supply, thermal dissipation, and
computing budget of onboard satellite platforms~\cite{ref:cuhkcr}.
Therefore, onboard cloud-removal models should not only pursue offline
PSNR/SSIM performance, but also consider inference energy, computation
cost, and deployment feasibility.

Spiking neural networks (SNNs) provide a computational paradigm
different from conventional ANNs for low-energy onboard intelligence.
SNNs transmit information through sparse event-driven spike activities,
allowing silent neurons to skip unnecessary computation and enabling
low-cost spike operations on neuromorphic hardware~\cite{ref:maass1997,
ref:loihi}. For LEO satellite platforms constrained by solar power,
payload energy consumption, and thermal control, such sparse
computation is highly attractive. Previous studies have explored the
potential of SNNs in low-level vision restoration tasks such as image
deraining~\cite{ref:esdnet}, and SNNs have also been applied to
decentralized federated classification in satellite
scenarios~\cite{ref:flsnn}. However, optical cloud removal imposes
stronger requirements on SNNs. Binary spikes can reduce energy
consumption but may lose the fine-grained activation representation
needed for pixel-level regression. Meanwhile, cloud occlusion is
spatially non-uniform and spectrally complex, requiring the model to
capture cloud boundaries, semi-transparent thin-cloud structures, and
high-frequency ground textures simultaneously. Thus, an onboard SNN
cloud-removal model must balance sparse low-energy computation with
high-fidelity restoration.

In addition to inference energy, onboard model training is also
challenging. Due to differences in orbital positions, observation
regions, sensor states, atmospheric conditions, and cloud types, local
data on different satellites usually exhibit significant source-level
heterogeneity. For example, some satellites may observe more thin-cloud
samples, whereas others may mainly encounter thick clouds or complex
cloud distributions. Centralizing all raw observations on the ground
for training would incur high communication overhead and also
contradict the trend toward autonomous in-orbit processing emphasized
by Space-CPN. Federated learning provides a natural
solution~\cite{ref:fedavg}: each satellite trains a model on its local
data and exchanges model parameters rather than raw observations.
However, standard centralized federated learning typically relies on a
ground parameter server, whereas LEO constellations are characterized
by intermittent connectivity and constrained intra-/inter-plane
communication topologies. Decentralized inter-satellite collaborative
training~\cite{ref:gossip-boyd} is therefore more suitable for
constellation scenarios. Moreover, source-level heterogeneous data
induce feature distribution shifts across satellites, and globally
shared batch-normalization statistics may destroy local feature
calibration. FedBN preserves local BN statistics to mitigate
feature-shift non-IID, which matches the source-level heterogeneous
satellite cloud-removal setting considered in this
work~\cite{ref:fedbn}.

The closest related work is FLSNN~\cite{ref:flsnn}, which demonstrates
the feasibility of decentralized federated SNN training on a
Walker-Star constellation for the EuroSAT classification task. However,
directly applying this paradigm to onboard optical cloud removal faces
three key gaps. The first is the \emph{task gap}: classification
produces discrete labels, whereas cloud removal is dense pixel
regression with different gradient structures, loss functions, and
restoration objectives. The second is the \emph{model gap}: satellite
cloud removal requires the reconstruction of multi-scale textures,
cloud boundaries, and spectral details, while small spiking
classification networks are insufficient for high-fidelity
reconstruction. The third is the \emph{data-heterogeneity gap}: non-IID
satellite observations in practical cloud-removal scenarios are driven
more by source factors such as sensors, cloud thickness, atmospheric
conditions, and observation regions than by class-label imbalance.
These gaps indicate that onboard federated cloud removal requires a
co-design across the task, model, and system levels.

To address these challenges, this paper studies a new problem:
\emph{decentralized onboard federated optical cloud removal over
satellite constellations}. To the best of our knowledge, this is the
first work to formulate optical cloud removal under this setting.
Different from existing ground-centered cloud-removal methods, this
work focuses on remote-sensing restoration under onboard communication,
energy, and computation constraints. Different from existing onboard
federated learning studies, this work addresses high-fidelity dense
image restoration rather than relatively lightweight tasks such as
classification or compression. The goal is not merely to improve
offline ground-side restoration metrics, but to establish a unified
design framework that jointly considers cloud-removal quality,
inference energy, onboard computation, and decentralized training.

To this end, we propose \textbf{\OrbitALIF{}}, an
\emph{Orbital Attention-enhanced Leaky Integrate-and-Fire} framework
for onboard optical cloud removal. \OrbitALIF{} exploits the sparse
event-driven computation of SNNs to reduce inference energy, while
improving pixel-level restoration capability through attention- and
receptive-field-enhanced modules and architecture. At the model level,
\OrbitALIF{} introduces five-level quantized spikes to improve the
representational capability of conventional binary spikes, and combines
dual-frequency residual modeling, frequency-spatial attention, adaptive
gated fusion, and spectro-temporal spike enhancement to recover cloud
boundaries, thin-cloud textures, and high-frequency ground details. At
the system level, \OrbitALIF{} constructs a decentralized constellation
training pipeline, retaining local BN statistics~\cite{ref:fedbn} to
mitigate source-level feature shift and exchanging shared parameters
through intra-plane aggregation and inter-plane Gossip
mixing~\cite{ref:gossip-boyd}. FedBN and Gossip are not proposed here
as new federated optimization algorithms; rather, they are adopted as
system-level training strategies adapted to constellation communication
topology and source-level heterogeneous data.

The main contributions of this work are summarized as follows.

\begin{itemize}
\item \textbf{We formulate decentralized onboard federated optical
  cloud removal.} To the best of our knowledge, this work is the first
  to formulate optical cloud removal as a decentralized onboard
  federated learning problem over satellite constellations, unifying
  optical remote-sensing restoration, onboard intelligent processing,
  and inter-satellite collaborative learning in a single framework.

\item \textbf{We propose \OrbitALIF{} for low-energy onboard cloud
  removal.} \OrbitALIF{} combines attention-enhanced LIF dynamics with
  sparse spiking computation. Through five-level quantized spikes,
  dual-frequency residual modeling, frequency-spatial attention,
  adaptive gated fusion, and spectro-temporal spike enhancement,
  \OrbitALIF{} preserves pixel-level restoration capability under
  low-power computation.

\item \textbf{We design a decentralized constellation training pipeline
  for source-level heterogeneous data.} To handle distribution
  differences across satellite observations, \OrbitALIF{} retains local
  BN statistics to mitigate feature shift and exchanges shared
  parameters through intra-plane aggregation and inter-plane Gossip
  mixing, avoiding both large-scale raw-data downlink and dependence on
  a centralized ground parameter server.

\item \textbf{We validate the quality--energy advantage and cross-task
  restoration potential of \OrbitALIF{}.} Experiments show that the
  $2.30$\,M onboard deployment variant improves cloud-removal quality
  while reducing per-image inference energy by up to $72.3\!\times$.
  The $14$\,M scaled variant reaches SOTA performance on image
  deraining and achieves competitive near-SOTA performance on optical
  cloud removal, indicating that attention-enhanced LIF dynamics have
  broader potential for low-level vision restoration.
\end{itemize}

The remainder of this paper is organized as follows.
Section~\ref{sec:sysmodel} introduces the constellation model, the
federated learning formulation, and the source-level Dirichlet non-IID
partition. Section~\ref{sec:framework} presents the \OrbitALIF{}
architecture and decentralized training pipeline.
Section~\ref{sec:experiments} reports centralized cloud-removal
results, component ablations, federated training results, and energy
analysis. Section~\ref{sec:conclusion} concludes this paper.

% =====================================================================
% Section II / III / IV / V — TO BE COMPOSED BY THE USER
% =====================================================================
% =====================================================================
% Section II: System Model
% =====================================================================
\section{System Model}\label{sec:sysmodel}

\subsection{LEO Constellation and Inter-Satellite Links}\label{sec:sysmodel-constellation}

We consider a Walker-Star low-Earth-orbit (LEO) satellite constellation
denoted by
\begin{equation}
\mathcal{S}
\;=\;
\left\{ s_{i,k} : i \in \mathcal{N}, \; k \in \mathcal{K}_i \right\},
\label{eq:constellation}
\end{equation}
where $\mathcal{N}$ is the set of orbital planes, $N\!=\!|\mathcal{N}|$
is the number of planes, and $\mathcal{K}_i$ is the set of satellites
in the $i$-th plane.  For notational simplicity we assume each plane
contains the same number of satellites,
$|\mathcal{K}_i|\!=\!K$ for all $i\!\in\!\mathcal{N}$.
Each satellite $s_{i,k}$ is equipped with inter-satellite laser links
(ISLs) for in-constellation model exchange. According to the
connection pattern, ISLs are divided into two categories: intra-plane
links and inter-plane links.

Intra-plane links connect adjacent satellites within the same orbital
plane. Since satellites in the same plane have relatively stable
geometry, these links are quasi-stationary across orbital periods and
are suitable for reliable ring-style aggregation. Inter-plane links
connect satellites from different orbital planes. Their availability
is more time-varying because of relative orbital motion, visibility
constraints, and Doppler effects. Following the Walker-Star setting
adopted in prior decentralized satellite
learning~\cite{ref:flsnn}, we instantiate $N\!=\!5$ and $K\!=\!10$,
yielding $NK\!=\!50$ satellites in total. For inter-plane
communication, we adopt a fixed chain topology as a simplified mixing
graph.

\subsection{Federated Optical Cloud-Removal Problem}\label{sec:sysmodel-problem}

Each satellite $s_{i,k}$ holds a local paired image dataset
\begin{equation}
\mathcal{D}_{i,k}
\;=\;
\left\{ (c_n, y_n) \right\}_{n=1}^{|\mathcal{D}_{i,k}|},
\label{eq:dataset}
\end{equation}
where $c_n\!\in\!\mathbb{R}^{3 \times H \times W}$ denotes a cloudy
image tile and $y_n\!\in\!\mathbb{R}^{3 \times H \times W}$ denotes
the corresponding cloud-free reference tile.  Let
$g_x : \mathbb{R}^{3 \times H \times W} \!\rightarrow\!
\mathbb{R}^{3 \times H \times W}$ be the \OrbitALIF{} cloud-removal
model parameterized by $x \!\in\! \mathbb{R}^d$, where
$d \!\approx\! 2.30 \!\times\! 10^6$ for the onboard deployment
variant. Let $\ell(g_x(c), y)$ denote the pixel-level reconstruction
loss, defined in Section~\ref{sec:framework-loss}. The global training
objective over the whole constellation is to minimise the empirical
risk
\begin{equation}
\min_{x \in \mathbb{R}^d}
f(x)
\;=\;
\frac{1}{NK}
\sum_{i \in \mathcal{N}}
\sum_{k \in \mathcal{K}_i}
f_{i,k}(x),
\label{eq:globalobj}
\end{equation}
where the local empirical risk on satellite $s_{i,k}$ is
\begin{equation}
f_{i,k}(x)
\;=\;
\frac{1}{|\mathcal{D}_{i,k}|}
\sum_{(c, y) \in \mathcal{D}_{i,k}}
\ell\!\left( g_x(c), \, y \right).
\label{eq:localobj}
\end{equation}

A centralised solution would require all paired cloudy/clear image
tiles to be downlinked to ground servers for training. This
contradicts the bandwidth and latency constraints discussed in
Section~\ref{sec:intro}. We therefore adopt a fully onboard
decentralised training strategy, where satellites compute local
updates on their own data and exchange model parameters through ISLs
without uploading raw observations.

\subsection{Decentralised Onboard Training Procedure}\label{sec:sysmodel-procedure}

The training proceeds in $T$ global communication rounds. Round $0$
starts from a common random initialisation $x^{(0)}\!\in\!\mathbb{R}^d$
that is broadcast to every satellite once at deployment time. Each
subsequent round $t\!\in\!\{0, \ldots, T-1\}$ then consists of four
steps.

\subsubsection*{S1: Local update}
At the beginning of round $t$, each satellite initialises its model
from its plane-level broadcast model:
\begin{equation*}
x_{i,k}^{(t, 0)} \;=\; x_i^{(t)},
\end{equation*}
with $x_i^{(0)}\!=\!x^{(0)}$ at round $t\!=\!0$.
Satellite $s_{i,k}$ then performs $E$ local AdamW steps on its dataset
$\mathcal{D}_{i,k}$:
\begin{equation}
x_{i,k}^{(t, e+1)}
\;=\;
\mathrm{AdamW}\!\left(
x_{i,k}^{(t, e)},
\,
\nabla \tilde{f}_{i,k}\!\left( x_{i,k}^{(t, e)} \right),
\,
\eta_t
\right),
\quad e\!=\!0, \ldots, E\!-\!1,
\label{eq:localstep}
\end{equation}
where $\nabla\tilde{f}_{i,k}$ is the mini-batch stochastic estimate
of $\nabla f_{i,k}$ and $\eta_t$ is the learning rate at round $t$.
The notation $\mathrm{AdamW}(\cdot,\cdot,\cdot)$ stands for the
abstract local-update operator implemented by the AdamW optimiser;
replacing it by SGD or any other optimiser leaves the rest of the
training procedure unchanged.

\subsubsection*{S2: Intra-plane aggregation}
After local updates, the $K$ satellites in each orbital plane perform
ring AllReduce over quasi-stationary intra-plane ISLs. This produces
the plane-level consensus model
\begin{equation}
x_i^{(t + \frac{1}{2})}
\;=\;
\frac{1}{K}
\sum_{k \in \mathcal{K}_i}
x_{i,k}^{(t, E)}.
\label{eq:intraplane}
\end{equation}
To accommodate source-level feature heterogeneity, we adopt the
\textsc{FedBN}~\cite{ref:fedbn} normalisation protocol: only
batch-normalisation \emph{affine} parameters $(\gamma, \beta)$ are
included in $x_{i,k}^{(t,E)}$ and therefore enter the average in
(\ref{eq:intraplane}); the BN \emph{running statistics}
(running mean and running variance) are kept per-satellite and are
not exchanged through either ring AllReduce or, later, the inter-plane
Gossip step.

\subsubsection*{S3: Inter-plane Gossip mixing}
After intra-plane aggregation, each orbital plane obtains a plane-level
model $x_i^{(t+\frac{1}{2})}$. The $N$ plane-level models are then
mixed through single-step Gossip averaging~\cite{ref:gossip-boyd} over
the inter-plane chain topology:
\begin{equation}
x_i^{(t+1)}
\;=\;
\sum_{j \in \mathcal{N}}
W_{ij} \, x_j^{(t + \frac{1}{2})},
\quad i \in \mathcal{N},
\label{eq:gossip}
\end{equation}
where $W \!\in\! \mathbb{R}^{N \times N}$ is a symmetric doubly
stochastic mixing matrix satisfying
\begin{equation*}
W \mathbf{1} \!=\! \mathbf{1}, \quad
\mathbf{1}^{\!\top} W \!=\! \mathbf{1}^{\!\top}, \quad
W \!=\! W^{\!\top}.
\end{equation*}
Consistent with the \textsc{FedBN} convention adopted in S2, the
Gossip step is applied only to shared restoration parameters
(convolutional weights, BN affine parameters, and other dense layer
weights). The BN running statistics remain local. In this way, the
constellation shares restoration knowledge across orbital planes while
avoiding the destructive averaging of feature statistics that would
arise under pure FedAvg-style aggregation when satellites observe
different cloud thicknesses, sensors, and atmospheric conditions.

\subsubsection*{S4: Intra-plane broadcast}
Finally, each orbital plane broadcasts the mixed model
$x_i^{(t+1)}$ to all satellites $s_{i,k}\!\in\!\mathcal{K}_i$ in the
same plane. This model is used as the initialisation for the next
communication round. The above four steps complete one round of
decentralised onboard training.

\subsection{Source-Level Dirichlet Non-IID Partition}\label{sec:sysmodel-dirichlet}

To simulate the natural source-level feature heterogeneity of
satellite imagery, we do not use the conventional label-level Dirichlet
partition commonly adopted in classification tasks. Instead, we
construct a source-level non-IID partition over the training pool.
Specifically, the joint training set is divided into $S\!=\!2$ source
domains: CUHK-CR1 represents thin-cloud scenes, and CUHK-CR2
represents thick-cloud scenes~\cite{ref:cuhkcr}. For each satellite
$s_{i,k}$, we draw a source mixing vector
\begin{equation}
\mathbf{p}_{i,k}
\;=\;
\left( p_{i,k}^{(1)}, \, p_{i,k}^{(2)} \right)
\;\sim\;
\Dir\!\left( \alpha \mathbf{1}_S \right).
\label{eq:dirichlet}
\end{equation}
Satellite $s_{i,k}$ then samples image tiles from the $s$-th source
domain in proportion to $p_{i,k}^{(s)}$. To ensure executable local
training, each satellite is assigned at least five training samples.

The Dirichlet concentration parameter $\alpha$ controls the strength
of source-level heterogeneity. A smaller $\alpha$ produces more
concentrated single-source clients, where some satellites mainly hold
thin-cloud samples while others mainly hold thick-cloud samples. A
larger $\alpha$ yields a more balanced source mixture and approaches
an IID-like setting. We use $\alpha\!=\!0.1$ as the default
configuration. Under this setting, $30$ out of $50$ satellites hold
only the minimum of five samples ($60\%$ of the constellation), while
three satellites individually hold more than $100$ samples, with the
actual partition log recording the three largest client sizes as
$\{117, 118, 118\}$. This partition reflects both the data sparsity
and source-level heterogeneity that may arise in onboard cloud-removal
scenarios. Sensitivity to $\alpha$ is reported in
Section~\ref{sec:exp-federated}.

% Section III: Framework
\section{The \OrbitALIF{} Framework}\label{sec:framework}
\textit{[Section to be written by the user.]}

% Section IV: Experiments
\section{Experiments}\label{sec:experiments}
\textit{[Section to be written by the user.]}

% Section V: Conclusion
\section{Conclusion}\label{sec:conclusion}
\textit{[Section to be written by the user.]}

% =====================================================================
% Bibliography
% =====================================================================
% IMPORTANT: Each \bibitem is tagged with one of three statuses:
%   [VERIFIED]    — bibliographic fields are stable across multiple
%                   independent sources (canonical / well-cited paper).
%                   Safe to keep as-is.
%   [USER VERIFY] — the paper exists in the public record but specific
%                   fields (page numbers, volume, doi, exact title
%                   capitalisation) should be cross-checked against the
%                   actual published / arXiv PDF before submission.
%   [PLACEHOLDER] — bibliographic skeleton only; user MUST replace with
%                   a real, verified citation.  Do not submit while any
%                   [PLACEHOLDER] entry remains.
% =====================================================================

\begin{thebibliography}{99}

% [PLACEHOLDER]
% Citation context: opening sentence, "real-time applications such as
% natural disaster monitoring, emergency response, agricultural
% observation...".  Recommendation: replace with a recent IEEE
% Geoscience and Remote Sensing Magazine survey on optical remote-
% sensing applications.  Sample candidates the user may consider
% (must be cross-checked individually):
%   - "Deep learning in remote sensing: A comprehensive review and
%      list of resources", Zhu et al., IEEE GRSM 2017.
%   - "Deep Learning Methods for Optical Remote Sensing Image
%      Restoration: A Survey", Liu et al., IEEE GRSM (year varies).
\bibitem{ref:earthobs1}
[PLACEHOLDER -- replace with verified survey on remote-sensing
applications.]

% [PLACEHOLDER]
% Citation context: same opening sentence, second authority citation.
% Recommendation: a survey on real-time satellite imagery for disaster
% / emergency response (e.g., IEEE GRSM or ISPRS Journal).
\bibitem{ref:earthobs2}
[PLACEHOLDER -- replace with verified survey on real-time satellite
imagery for disaster / emergency response.]

% [USER VERIFY]
% Title and venue match the public CVPRW 2024 (AI4Space workshop)
% paper "Tackling the Satellite Downlink Bottleneck with Federated
% Onboard Learning of Image Compression" by Gomez et al.  The CVF
% open-access page is at openaccess.thecvf.com (CVPRW 2024 AI4Space).
% USER ACTION: open the CVF PDF and confirm authors / page range /
% exact title before submission.
\bibitem{ref:onboardfl-bottleneck}
P.~Gomez \emph{et~al.},
``Tackling the satellite downlink bottleneck with federated onboard
learning of image compression,''
\emph{Proc.\ IEEE/CVF Conf.\ Comput.\ Vis.\ Pattern Recognit.\
Workshops (CVPRW)}, 2024.

% [PLACEHOLDER]
% Citation context: Space-CPN concept and architecture.  Recent
% Chinese-language and IEEE Network articles discuss this concept.
% USER ACTION: replace with one or two specific Space-CPN architecture
% papers (IEEE Network / IEEE Comm.\ Magazine).  Make sure the cited
% paper actually defines or surveys Space Computing Power Networks.
\bibitem{ref:spacecpn1}
[PLACEHOLDER -- replace with verified Space Computing Power Networks
architecture / survey paper.]

% [USER VERIFY]
% This is the CUHK-CR / DDFR paper.  An arXiv version is available at
% arXiv:2401.15105 with title
%   "Diffusion Enhancement for Cloud Removal in Ultra-Resolution
%    Remote Sensing Imagery"
% and the published version is in IEEE TGRS.  USER ACTION: confirm
% the final IEEE TGRS volume / issue / pages and the canonical author
% list against the arXiv abstract page or IEEE Xplore record.
\bibitem{ref:cuhkcr}
Y.~Sui \emph{et~al.},
``Diffusion enhancement for cloud removal in ultra-resolution remote
sensing imagery,''
\emph{IEEE Trans.\ Geosci.\ Remote Sens.}, 2024.
% arXiv:2401.15105

% [VERIFIED]
% Canonical SNN reference establishing spiking neurons as the third
% generation of neural network models.  Bibliographic fields are
% stable across decades of citations.
\bibitem{ref:maass1997}
W.~Maass,
``Networks of spiking neurons: The third generation of neural network
models,''
\emph{Neural Networks}, vol.~10, no.~9, pp.~1659--1671, 1997.

% [USER VERIFY]
% Loihi neuromorphic processor.  USER ACTION: confirm exact
% bibliographic fields (volume / issue / month) of IEEE Micro.
\bibitem{ref:loihi}
M.~Davies \emph{et~al.},
``Loihi: A neuromorphic manycore processor with on-chip learning,''
\emph{IEEE Micro}, vol.~38, no.~1, pp.~82--99, Jan./Feb.\ 2018.

% [USER VERIFY]
% ESDNet for SNN-based image deraining; arXiv preprint listed at
% arXiv:2207.02094.  USER ACTION: confirm whether a peer-reviewed
% version exists in a specific venue and update accordingly.
\bibitem{ref:esdnet}
H.~Song \emph{et~al.},
``Exploring the potentials of spiking neural networks for image
deraining,''
\emph{arXiv preprint arXiv:2207.02094}, 2022.

% [USER VERIFY]
% FLSNN: decentralized federated SNN over Walker-Star constellation
% for EuroSAT classification.  arXiv preprint identifier listed in
% the user's existing draft as arXiv:2501.15995, 2025.  USER ACTION:
% open the arXiv abstract page, copy the exact title and author list,
% and confirm the preprint ID.  This is a load-bearing citation for
% the related-work paragraph; treat it as critical.
\bibitem{ref:flsnn}
P.~Yang, T.~Wang, \emph{et~al.},
``Brain-inspired decentralized satellite learning in space computing
power networks,''
\emph{arXiv preprint arXiv:2501.15995}, 2025.

% [VERIFIED]
% FedAvg.  Canonical federated learning baseline; bibliographic fields
% are widely consistent.
\bibitem{ref:fedavg}
H.~B.~McMahan, E.~Moore, D.~Ramage, S.~Hampson, and B.~A.~y~Arcas,
``Communication-efficient learning of deep networks from decentralized
data,''
\emph{Proc.\ Int.\ Conf.\ Artif.\ Intell.\ Statist.\ (AISTATS)}, 2017.

% [VERIFIED]
% FedBN.  ICLR 2021.  Title and authorship are widely cross-cited.
\bibitem{ref:fedbn}
X.~Li, M.~Jiang, X.~Zhang, M.~Kamp, and Q.~Dou,
``FedBN: Federated learning on non-IID features via local batch
normalization,''
\emph{Proc.\ Int.\ Conf.\ Learn.\ Represent.\ (ICLR)}, 2021.

% [VERIFIED]
% Boyd et al. on randomized gossip algorithms.  Canonical reference
% for gossip-style decentralized averaging.
\bibitem{ref:gossip-boyd}
S.~Boyd, A.~Ghosh, B.~Prabhakar, and D.~Shah,
``Randomized gossip algorithms,''
\emph{IEEE Trans.\ Inf.\ Theory}, vol.~52, no.~6, pp.~2508--2530,
Jun.\ 2006.

\end{thebibliography}

\end{document}
